---
title: "Metaanalysis of COVID-19 incidence velocities"
author: 
  - "Vincent J. Carey, stvjc at channing.harvard.edu"
  - "Sean Davis, seandavi at gmail.com"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Metaanalysis of COVID-19 incidence velocities}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

The `Arima*` functions of sars2app model COVID-19 incidence series.
Using ARIMA(0,1,3) by default we get an estimate of drift and the
standard error of this estimate.

For the US as a whole, as of May 17 2020:
```{r lkd1, cache=TRUE}
library(sars2app)
ej = enriched_jhu_data()
usa_full = Arima_nation(ej)
usa_full
```

For the US dropping NY:
```{r lkd2, cache=TRUE}
nyd = nytimes_state_data()
drny = Arima_drop_state(ej, nyd)
drny

```{r lkd, fig.width=11, fig.height=5,cache=TRUE}
par(mfrow=c(1,2), mar=c(4,3,2,2))
plot(usa_full, main="USA aggregated", ylim=c(15900,36000))
plot(drny, main="USA excluding NY", ylim=c(15900,36000) )
```

Two different data sources are used for this display, as
certain anomalies were apparent on May 17 2020 that
inhibited summing NYT data to get US aggregate.  The
reconciliation of NYT, JHU and USA facts daily data could be
of interest.

# Infection velocity estimates for contiguous 48 states and DC

```{r doall,cache=TRUE}
allst = contig_states_dc()
allarima = lapply(allst, function(x) Arima_by_state(nyd, x))
names(allarima) = allst
drifts = sapply(allarima, function(x) coef(x$fit)["drift"])
searima = function(a) sqrt(a$fit$var.coef["drift", "drift"])
se.drifts = sapply(allarima, searima)
```

```{r lkmeta}
library(rmeta)
meta.summaries(drifts, se.drifts)
```{r dorm, fig.height=10}
o = order(drifts)
metaplot(drifts[o], se.drifts[o], labels=allst[o], cex=.7, 
  xlab="Infection velocity (CHANGE in number of confirmed cases/day)", ylab="State")
segments(rep(-350,46), seq(-49,-4), rep(-50,46), seq(-49,-4), lty=3, col="gray")
```

